Phylogenetics and the human microbiome

### 2013-04-02 sbreview.tex
lots more citations

@@

@@

@@

@@

@@

UniFrac distance and variants.

@@

@@

@@

@@

There has been a shift from interest in community dynamics to fine-scale interactions between host and immune system.

\cite{wylie2012sequence}
\cite{chen2012associating}

@@

>>>>

\cite{hugenholtz2002review}
{{E}xploring prokaryotic diversity in the genomic era}

@@

Core microbiome
\cite{turnbaugh2008core}

HMP nature paper
\cite{methe2012framework}

% applications
\cite{costello2009bacterial}
Bacterial community variation in human body habitats across space and time

\cite{ley2006microbial}
Microbial ecology: human gut microbes associated with obesity

@@

\cite{parks2011classifying}
Classifying short genomic fragments from novel lineages using composition and homology

\cite{mcdonald2011improved}
An improved Greengenes taxonomy with explicit ranks for ecological and evolutionary analyses of bacteria and archaea

\cite{phylopythia}
McHardy, A. C.  and Martin, H. G.  and Tsirigos, A.  and Hugenholtz, P.  and Rigoutsos, I.

\cite{munch2008statistical}
Statistical assignment of DNA sequences using Bayesian phylogenetics

\cite{munch2008fast}
Fast phylogenetic DNA barcoding

\cite{rosen2011nbc}
NBC: the Naive Bayes Classification tool webserver for taxonomic classification of metagenomic reads

\cite{segata2011metagenomic}
Metagenomic biomarker discovery and explanation

\cite{segata2012metagenomic}
Metagenomic microbial community profiling using unique clade-specific marker genes

@@

\cite{forey2001phylocode}
{The PhyloCode: description and commentary}

\cite{matsen2011reconciling}
Reconciling taxonomy and phylogenetic inference: formalism and algorithms for describing discord and inferring taxonomic roots

\cite{carma}

\cite{kreig1984bergey}
Bergey's Manual of Systematic Bacteriology.

\cite{dalevi2007automated}
Automated group assignment in large phylogenetic trees using GRUNT: GRouping, Ungrouping, Naming Tool

\cite{desantis2006greengenes}
{Greengenes, a chimera-checked 16S rRNA gene database and workbench compatible with ARB}

\cite{bazinet2012comparative}
A comparative evaluation of sequence classification programs

\cite{brady2011phymmbl}
PhymmBL expanded: confidence scores, custom databases, parallelization and more

\cite{brady2009phymm}
Phymm and PhymmBL: metagenomic phylogenetic classification with interpolated Markov models

\cite{ColeEaRDP2009}

\cite{liu2008accurate}
Accurate taxonomy assignments from 16S rRNA sequences produced by highly parallel pyrosequencers

\cite{lan2012using}
Using the RDP Classifier to Predict Taxonomic Novelty and Reduce the Search Space for Finding Novel Organisms

\cite{macdonald2012rapid}
Rapid identification of high-confidence taxonomic assignments for metagenomic data

\cite{maidak1997rdp}
The RDP (ribosomal database project)

@@

\cite{hill1973diversity}
Diversity and evenness: a unifying notation and its consequences

\cite{odwyer2012phylogenetic}
Phylogenetic Diversity Theory Sheds Light on the Structure of Microbial Communities

cite kembel paper

\cite{morlon2011spatial}
Spatial patterns of phylogenetic diversity

@@

\cite{chen2012associating}
Associating microbiome composition with environment-- generalized unifrac

UniFrac distance and variants.
\cite{LozuponeKnightUniFrac05}
\cite{LozuponeEaWeightedUnifrac07}

@@

\cite{BikEaMicrobiotaStomach06}

hill numbers
\cite{chao2010phylogenetic}

\cite{evans2012phylogenetic}
The phylogenetic Kantorovich-Rubinstein metric for environmental sequence samples

\cite{kuczynski2010microbial}
Microbial community resemblance methods differ in their ability to detect biologically relevant patterns

\cite{jaccard1908nouvelles}
Nouvelles recherches sur la distribution florale

\cite{matsen2013edge}
Edge principal components and squash clustering: using the special structure of phylogenetic placement data for sample comparison

\cite{PurdomAnalyzingDataGraphs08}

\cite{nemergut2011global}
?? Global patterns in the biogeography of bacterial taxa

@@

\cite{felsenstein1981evolutionary}
{Evolutionary trees from DNA sequences: a maximum likelihood approach}

\cite{eddy1998profile}
Profile hidden Markov models.

\cite{li2006cdhit}
Li, W.  and Godzik, A.
\cite{edgar2010usearch}

\cite{mirarabsepp}
{SEPP: SAT{\'e}-Enabled Phylogenetic Placement}

\cite{monierEaLargeViruses08}
{Taxonomic distribution of large DNA viruses in the sea}

\cite{huson2007megan}

\cite{mcnabb2004hsp65}
{{A}ssessment of partial sequencing of the 65-kilodalton heat shock protein gene (hsp65) for routine identification of {M}ycobacterium species isolated from clinical sources}

\cite{price2010fasttree}
{FastTree 2--approximately maximum-likelihood trees for large alignments}

\cite{navlakha2009finding}
Finding biologically accurate clusterings in hierarchical tree decompositions using the variation of information

\cite{nawrocki2009infernal}
Infernal 1.0: inference of RNA alignments

\cite{white2010alignment}
Alignment and clustering of phylogenetic markers-implications for microbial diversity studies

RNA copy number
\cite{morganEaInVitroSimulatedMetagenome10}

@@

\cite{wu2008simple}
A simple, fast, and accurate method of phylogenomic inference

\cite{matsen2010pplacer}

\cite{ludwig2004arb}
{{ARB}: a software environment for sequence data}
\cite{berger2011performance}

\cite{berger2011aligning}
Aligning short reads to reference alignments and trees

\cite{caporaso2010qiime}

\cite{schloss2009introducing}
Introducing mothur: open-source, platform-independen

\cite{lynch2013axiome}
AXIOME: automated exploration of microbial diversity

\cite{wang2007naive}
{Naive Bayesian classifier for rapid assignment of rRNA sequences into the new bacterial taxonomy}

\cite{werner2011impact}
Impact of training sets on classification of high-throughput bacterial 16s rRNA gene surveys

\cite{stark2010mltreemap}
{{MLTreeMap}-accurate Maximum Likelihood placement of environmental DNA sequences into taxonomic and functional reference phylogenies.}

\cite{vonMeringEaQuantitative08}
{Quantitative phylogenetic assessment of microbial communities in diverse environments}

% alternative loci
\cite{case2007rpob}

% databases
\cite{chen2010human}
The Human Oral Microbiome Database: a web accessible resource for investigating oral microbe taxonomic and genomic information

\cite{griffen2011core}
CORE: a phylogenetically-curated 16S rDNA database of the core oral microbiome

\cite{srinivasan2012bacterial}

@@

There has been a shift from interest in community dynamics to fine-scale interactions between host and immune system.
\cite{wylie2012sequence}
\cite{chen2012associating}

% medical
\cite{clarridge2004}
{{I}mpact of 16{S} r{RNA} gene sequence analysis for identification of bacteria on clinical microbiology and infectious diseases}

\cite{phillips2012microbiome}
Microbiome analysis among bats describes influences of host phylogeny, life history, physiology and geography

@@

\cite{hummelen2010deep}
Deep sequencing of the vaginal microbiota of women with HIV

\cite{ravel2011vaginal}
Vaginal microbiome of reproductive-age women

\cite{fredricks2005molecular}
Molecular identification of bacteria associated with bacterial vaginosis
